Document Classification

### 2018-11-03 DATA607/Project_4.Rmd
Finished project 4

## Description

This project seeks to build a classifier that can predict if an email is spam or ham.  It will be using the Apache Spam Assassin public corpus as the dataset.

### Data Aquisition

@@

```{r file_download}
if(!file.exists("data/20050311_spam_2.tar.bz2") && !file.exists("data/20050311_spam_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar.bz2")
}

if(!file.exists("data/20030228_easy_ham_2.tar.bz2") & !file.exists("data/20030228_easy_ham_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar.bz2")
}
```

Now that the files have been aquired, they need to be decompressed.

```{r decompress_files}
if(!dir.exists("data/spam_2")){
  R.utils::bunzip2("data/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar")
  untar("data/20050311_spam_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/spam_2/cmds")
}

if(!dir.exists("data/easy_ham_2")){
  R.utils::bunzip2("data/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar")
  untar("data/20030228_easy_ham_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/easy_ham_2/cmds")
}
```

## Processing Pipeline

### Train/Test Split

First we are going to split the data into training and testing sets using an 80:20 split.

```{r}
set.seed(12345)
training_percent <- 0.8

train_test_split <- function(data, training_percent){
  n <- round(length(data)*training_percent, digits = 0)
  train <- sort(sample(data, n, replace = FALSE))
  test <- data[!data %in% train]
  return(list(test=test, train=train))
}

spam <- train_test_split(list.files("data/spam_2"), training_percent)
ham <- train_test_split(list.files("data/easy_ham_2"), training_percent)
```

The following table summarizes the number of messages included in each set:

##### Table 1. Spam and Ham Message Counts by Data Set
|      |       Training       |        Testing      |
|------|----------------------|---------------------|
| Spam |`r length(spam$train)`|`r length(spam$test)`|
| Ham  |`r length(ham$train)` |`r length(ham$test)` |

### Pull out Email Contents

This model will classify only on the basis of the content.  The header information will not be taken into consideration.  In order to illustrate the process I will take the first message in the spam training set as an example.

```{r}
example_file <- spam$train[1]

get_email_contents <- function(file_name, path){
 full_path = paste0(path, file_name)
 con <- file(full_path)
 readLines(con)
}

```

### Split up Sentance

### Remove Punctuation and Stop Words

### Stemming

>>>>

## Introduction

This project seeks to build a classifier that can predict if an email is *spam* or *ham*.  It will be using the Apache Spam Assassin public corpus as the dataset.

## Data Acquisition

@@

```{r file_download}
if (!file.exists("data/20050311_spam_2.tar.bz2") && !file.exists("data/20050311_spam_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar.bz2")
}

if (!file.exists("data/20030228_easy_ham_2.tar.bz2") & !file.exists("data/20030228_easy_ham_2.tar")){
  download.file("https://spamassassin.apache.org/old/publiccorpus/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar.bz2")
}
```

Now that the files have been acquired, they need to be decompressed.

```{r decompress_files}
if (!dir.exists("data/spam_2")){
  R.utils::bunzip2("data/20050311_spam_2.tar.bz2", "data/20050311_spam_2.tar")
  untar("data/20050311_spam_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/spam_2/cmds")
  # Delete these messages b/c it breaks the pipeline
  unlink("data/spam_2/00706.5116018237368c3633823b2d24f8ac86")
  unlink("data/spam_2/00708.89f1f9108884517148fdbd744e18ec1e")
  unlink("data/spam_2/00737.af5f503fe444ae773bfeb4652d122349")
  unlink("data/spam_2/01125.46ca779f86e1dd0a03c3ffc67b57f55e")
  unlink("data/spam_2/01217.d5a1734ec521c1bd55270eca3ab4acd8")
}

if (!dir.exists("data/easy_ham_2")){
  R.utils::bunzip2("data/20030228_easy_ham_2.tar.bz2", "data/20030228_easy_ham_2.tar")
  untar("data/20030228_easy_ham_2.tar", exdir = "data")
  # Delete the cmds file
  unlink("data/easy_ham_2/cmds")
}
```

## Build a Corpus

Now that we have a dataset we need to construct a corpus.  I will use the *tm* package and *dplyr* to get the job done.

```{r build_corpus, message=FALSE}
library(dplyr)
library(tm)

get_corpus <- function(dir, label){
  corpus <- VCorpus(DirSource(dir)) %>%
    tm_map(PlainTextDocument)  %>% # Create plain text document
    tm_map(content_transformer(tolower)) %>% # Standardize case
    tm_map(removeWords, stopwords("SMART")) %>% # Remove stopwords
    tm_map(removePunctuation) %>% # Remove punctuation marks
    tm_map(removeNumbers) %>% # Remove numbers
    tm_map(stripWhitespace) %>% # Remove extra whitespace
    tm_map(stemDocument) # Stem the documents
  meta(corpus, "LABEL") <- label
  return(corpus)
}

corpus<- c(get_corpus("data/spam_2", "Spam"), get_corpus("data/easy_ham_2", "Ham"))
```

## Build a Document-Term Matrix

The next step is to construct a document term matrix.

```{r dtm}
dtm <- DocumentTermMatrix(corpus)
```

There are `r dtm$nrow` documents and `r dtm$ncol` terms in the document-term matrix.

### Remove Infrequent Terms

There are a lot of terms resulting in a very sparce matrix.  I am going to prune out the less frequent terms from the matrix.  The term must be found in at least 10 documents.

```{r pruning}
min_docs <- 10
dtm <- removeSparseTerms(dtm, 1 - (min_docs / length(corpus)))
```

There are still `r dtm$nrow` documents, but there are `r dtm$ncol` terms in the pruned document-term matrix.

## Normalizing the Data

Some messages have more words than others.  We are going to normalize the data and pull all the data together into a model dataset.

```{r normalize_the_data}
model_data <- as.matrix(dtm)
words <- rowSums(model_data)
model_data <- model_data / words
model_data <- data.frame(model_data)
model_data <- cbind(meta(corpus), model_data) %>%
  mutate(LABEL = as.factor(LABEL))
```

## Split Data into Training/Test Data Sets

We are now ready to break up the data into training and test sets.  We are going to use a 75%/25% split.  This decision was arbitrary.

```{r train_test_split, message=FALSE}
library(caret)
set.seed(12345)
in_training_set <- createDataPartition(model_data$LABEL, p = 0.75,  list = FALSE)
training_data <- model_data[in_training_set, ]
testing_data <- model_data[-in_training_set, ]
```

## Build a SVM Model

We will take out training data and produce a SVM model that predicts if a message is spam or ham.

```{r build_svm}
library(e1071)
model <- svm(LABEL ~ ., data = training_data)
```

## Test the Model

Now that we have trained a model we can test it out and see how it preforms.

```{r test_model}
predictions <- testing_data %>%
  select(-LABEL) %>%
  predict(model, .)
```

#### Table 1. Confusion Matrix
```{r confusion_matrix, echo=FALSE}
library(kableExtra)
table(predictions, testing_data$LABEL) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```

```{r errors, echo=FALSE}
errors <- table(predictions, testing_data$LABEL) %>%
  data.frame(.) %>%
  mutate(keep = ifelse(predictions == Var2, 0, 1)) %>%
  filter(keep == 1) %>%
  summarise(total = sum(Freq))
```

There were `r errors$total` messages that were misclassified out of `r length(predictions)` for a `r round((1 - (errors$total / length(predictions))) *100, 0)`% accuracy rate.   Not bad!
